---
title: "SoCo Homebrew All-Grain Recipe Scraper"
format: html
---

I want to scrape all of the [all grain recipes](https://socohomebrew.com/homebrew-recipe-kits/all-grain-recipe-kits/?page=1) from SoCo Homebrew. I'll walk you through how I want to approach this.

## Get the urls for individual recipe order pages

This is the URL that stats listing all the recipe pages:

[All Grain Recipe Kits](https://socohomebrew.com/homebrew-recipe-kits/all-grain-recipe-kits/)

On that page there are some images that link to groups of recipes. I don't want those pages. But after the dropdown that says "Featured Items" there starts a list of each recipe's order page. We'll want to collect the URL for each of those pages.

At the bottom of the page is a link to advance to the next page, with a URL like this: [Page 2](https://socohomebrew.com/homebrew-recipe-kits/all-grain-recipe-kits/?page=2). There are 15 such pages. We want the URLs for recipe listed on all 15 pages.

## Find and download the recipe PDF

Here is an [example url for the recipe order page](https://socohomebrew.com/imperial-bourbon-barrel-chocolate-stout-all-grain/).

One that page is a text leadin "Recipe Sheet Download" with hyperlinked text right after it. That text changes for each recipe, but it links to the PDF for that individual recipe.

For each recipe order page, I want that PDF downloaded into a folder called recipe-pdfs.

I want you to wait 2 seconds between each download.

Build the scraper using the rvest package in the page called 01-scrape.qmd

***

```{r}
#| label: setup
library(rvest)
dir.create("recipe-pdfs", showWarnings = FALSE)
```

## Step 1: Collect recipe page URLs from all 15 listing pages

```{r}
#| label: collect-urls

get_recipe_urls <- function(page_num) {
  url <- paste0(
    "https://socohomebrew.com/homebrew-recipe-kits/all-grain-recipe-kits/?page=",
    page_num
  )
  page <- read_html(url)
  page |>
    html_elements(".productGrid .product h4 a") |>
    html_attr("href")
}

recipe_urls <- lapply(1:15, get_recipe_urls) |> unlist()

cat("Total recipe pages found:", length(recipe_urls), "\n")
```

## Step 2: For each recipe page, find and download the PDF

```{r}
#| label: download-pdfs

get_pdf_url <- function(recipe_url) {
  page <- read_html(recipe_url)
  page |>
    html_elements("a") |>
    (\(links) {
      hrefs <- html_attr(links, "href")
      hrefs[grepl("\\.pdf$", hrefs, ignore.case = TRUE)]
    })() |>
    head(1)
}

download_recipe_pdf <- function(recipe_url) {
  pdf_url <- tryCatch(get_pdf_url(recipe_url), error = \(e) character(0))

  if (length(pdf_url) == 0 || is.na(pdf_url)) {
    message("No PDF found for: ", recipe_url)
    return(invisible(NULL))
  }

  filename <- utils::URLdecode(basename(pdf_url))
  dest_path <- file.path("recipe-pdfs", filename)

  if (file.exists(dest_path)) {
    message("Already downloaded: ", filename)
    return(invisible(dest_path))
  }

  tryCatch(
    {
      download.file(pdf_url, dest_path, mode = "wb", quiet = TRUE)
      message("Downloaded: ", filename)
    },
    error = \(e) message("Failed to download ", pdf_url, ": ", conditionMessage(e))
  )

  Sys.sleep(2)
  invisible(dest_path)
}

# Download all PDFs with a 2-second pause between each
for (url in recipe_urls) {
  download_recipe_pdf(url)
}
```

## Results

```{r}
#| label: results

pdfs <- list.files("recipe-pdfs", pattern = "\\.pdf$", ignore.case = TRUE)
cat("PDFs downloaded:", length(pdfs), "\n")
```
