---
title: "Extract Recipe Data from PDFs"
format: html
---

```{r}
#| label: setup
#| message: false

library(pdftools)
library(tidyverse)
```

## Helper functions

```{r}
#| label: helpers

# Reconstruct text from a horizontal slice of a PDF page (coordinate-based)
reconstruct_column <- function(page_data, x_min = 0, x_max = Inf) {
  page_data |>
    filter(x >= x_min, x < x_max) |>
    arrange(y, x) |>
    mutate(line_group = cumsum(c(0, diff(y)) > 3)) |>
    group_by(line_group) |>
    summarise(line = paste(text, collapse = " "), .groups = "drop") |>
    pull(line) |>
    paste(collapse = "\n")
}

# Some PDFs split "hops:" across lines as "ho\n  ps:" — rejoin before parsing
preprocess_col <- function(text) {
  str_replace_all(text, "(?m)ho\n\\s*ps:", "hops:\n")
}

# Pull lines that look like ingredient quantities (e.g. "9 LB Pale Malt")
extract_ingredients <- function(text) {
  lines <- str_split(text, "\n")[[1]]
  hits  <- lines[str_detect(lines, "^\\s*\\d+\\.?\\d*\\s+(oz|LB|lb|lbs|g|kg)\\s+\\w")]
  str_trim(hits)
}

# Extract bittering / flavor / aroma hops — line-by-line search, robust to
# word-order scrambling that can occur in two-column coordinate reconstruction
extract_hop_section <- function(col_text, hop_type) {
  col_text <- preprocess_col(col_text)
  lines    <- str_split(col_text, "\n")[[1]]
  # Header: line containing hop_type AND ("hops:" or ends with "ho" = split word)
  header_idx <- which(
    str_detect(str_to_lower(lines), hop_type) &
    (str_detect(lines, "hops:") | str_detect(str_trim(lines), "\\bho$"))
  )
  if (length(header_idx) == 0) return(NA_character_)
  idx       <- header_idx[1]
  remaining <- lines[seq(idx + 1, length(lines))]
  # Stop at next hop header or STEP
  stop_idx <- which(
    (str_detect(str_to_lower(remaining), "bittering|flavor|aroma") &
       (str_detect(remaining, "hops:") | str_detect(str_trim(remaining), "\\bho$"))) |
    str_detect(remaining, "STEP")
  )
  if (length(stop_idx) > 0) remaining <- remaining[seq_len(stop_idx[1] - 1)]
  ing <- extract_ingredients(paste(remaining, collapse = "\n"))
  if (length(ing) == 0) return(NA_character_)
  paste(ing, collapse = "\n")
}

# Extract yeast entries — lines come in pairs: manufacturer / style
extract_yeast <- function(left_p2) {
  section <- str_extract(left_p2, "(?s)Recommended Yeast:\\n(.+?)(?=\nIf you have wet)")
  if (is.na(section)) return(NA_character_)
  lines <- str_split(section, "\n")[[1]]
  lines <- str_trim(lines[!str_detect(lines, "^Recommended Yeast:|^\\s*$")])
  if (length(lines) >= 2 && length(lines) %% 2 == 0) {
    pairs <- paste(lines[c(TRUE, FALSE)], lines[c(FALSE, TRUE)])
  } else {
    pairs <- lines
  }
  paste(pairs, collapse = "\n")
}

# Extract dry hop ingredients; empty form fields appear as "Text90:" prefix
extract_dry_hops <- function(left_p2) {
  section <- str_extract(left_p2, "(?s)secondary fermenter at this point:\\n(.+?)(?=\nAfter you add)")
  if (is.na(section)) return(NA_character_)
  lines <- str_split(section, "\n")[[1]]
  # Strip "Text90: " prefix; if line then has content, keep it
  lines <- str_trim(str_remove(lines, "^Text90:\\s*"))
  lines <- lines[lines != "" & !str_detect(lines, "secondary fermenter at this point")]
  if (length(lines) == 0) return(NA_character_)
  paste(lines, collapse = "\n")
}

# Extract flavoring — return NA if "No Flavoring"
extract_flavoring <- function(right_p2) {
  section <- str_extract(right_p2, "(?s)Flavoring to add before bottling\\n(.+?)(?=\nOnce the beer|$)")
  if (is.na(section)) return(NA_character_)
  lines <- str_split(section, "\n")[[1]]
  lines <- str_trim(lines[!str_detect(lines, "Flavoring to add before bottling|^\\s*$")])
  lines <- lines[lines != ""]
  if (length(lines) == 0 || all(lines == "No Flavoring")) return(NA_character_)
  paste(lines[lines != "No Flavoring"], collapse = "\n")
}
```

## Main extraction functions

Two PDF formats exist in this collection: the standard step-by-step "Recipe
Instructions" layout and a condensed "Recipe Sheet" layout. The wrapper
`extract_recipe()` auto-detects which format to use.

```{r}
#| label: extract-instructions

# Standard step-by-step "Recipe Instructions" format
extract_recipe_instructions <- function(pdf_path) {
  pages    <- pdf_data(pdf_path)
  raw_text <- paste(pdf_text(pdf_path), collapse = "\n")

  page_w       <- max(pages[[1]]$x + pages[[1]]$width)
  col_boundary <- page_w / 2

  left_p1  <- reconstruct_column(pages[[1]], x_min = 0,            x_max = col_boundary)
  right_p1 <- reconstruct_column(pages[[1]], x_min = col_boundary, x_max = Inf)
  left_p2  <- if (length(pages) >= 2) reconstruct_column(pages[[2]], x_min = 0,            x_max = col_boundary) else ""
  right_p2 <- if (length(pages) >= 2) reconstruct_column(pages[[2]], x_min = col_boundary, x_max = Inf) else ""

  # 1. Recipe name (parentheses allowed for e.g. "Session IPA (SIPA)")
  name_m      <- str_match(raw_text, "(?m)^\\s+([\\w][\\w\\s&''\\-()]+?)\\s{5,}Kit Type:")
  recipe_name <- if (!is.na(name_m[1, 2])) str_trim(name_m[1, 2]) else NA_character_

  # 2. Grains (left column, page 1)
  grain_section <- str_extract(left_p1, "(?s)following grain:\\n(.+?)(?=\n\nHeat|\nHeat)")
  grains_str    <- if (!is.na(grain_section)) {
    ing <- extract_ingredients(grain_section)
    if (length(ing) > 0) paste(ing, collapse = "\n") else NA_character_
  } else NA_character_

  # 3–5. Hops (right column, page 1)
  bittering_hops <- extract_hop_section(right_p1, "bittering")
  flavor_hops    <- extract_hop_section(right_p1, "flavor")
  aroma_hops     <- extract_hop_section(right_p1, "aroma")

  # 6 & 9. Specific gravity — use raw text to avoid coordinate-sort issues
  og_m        <- str_match(raw_text, "original specific gravity should be approximately:\\s*([\\d.]+)")
  original_sg <- if (!is.na(og_m[1, 2])) og_m[1, 2] else NA_character_
  fg_m        <- str_match(raw_text, "final specific gravity should be approximately:\\s*([\\d.]+)")
  final_sg    <- if (!is.na(fg_m[1, 2])) fg_m[1, 2] else NA_character_

  # 7. Yeast (left column, page 2)
  yeast_str <- extract_yeast(left_p2)

  # 8. Dry hops (left column, page 2)
  dry_hops_str <- extract_dry_hops(left_p2)

  # 10. Flavoring (right column, page 2)
  flavoring_str <- extract_flavoring(right_p2)

  tibble(
    recipe_name    = recipe_name,
    grains         = grains_str,
    bittering_hops = bittering_hops,
    flavor_hops    = flavor_hops,
    aroma_hops     = aroma_hops,
    original_sg    = original_sg,
    yeast          = yeast_str,
    dry_hops       = dry_hops_str,
    final_sg       = final_sg,
    flavoring      = flavoring_str
  )
}
```

```{r}
#| label: extract-sheet

# Condensed "Recipe Sheet" format (also used by some mislabeled "Instructions" files)
extract_recipe_sheet <- function(pdf_path) {
  raw_text <- paste(pdf_text(pdf_path), collapse = "\n")

  # 1. Recipe name — two variants:
  #    a) "Recipe Name: Foo Bar"
  #    b) Name on its own line before "Recipe Type:"
  name_m <- str_match(raw_text, "Recipe Name:\\s*(.+)")
  if (!is.na(name_m[1, 2])) {
    recipe_name <- str_trim(name_m[1, 2])
  } else {
    name_m2 <- str_match(raw_text, "(?m)^([A-Z][A-Za-z0-9 &'()\\-]+)\\s*\nRecipe Type:")
    recipe_name <- if (!is.na(name_m2[1, 2])) str_trim(name_m2[1, 2]) else NA_character_
  }

  # 2. Grains (header may have trailing right-column text on same line)
  grain_section <- str_extract(raw_text,
    "(?s)Grains Included in this Kit:[^\n]*\n(.+?)(?=\n[A-Z][a-z]+ [A-Z]|\\nExtract|\\nHops)")
  grains_str <- if (!is.na(grain_section)) {
    ing <- extract_ingredients(grain_section)
    if (length(ing) > 0) paste(ing, collapse = "\n") else NA_character_
  } else NA_character_

  # 3-5. Hops — three formats:
  #   a) "1 oz Foo Hops Boil for 60 minutes"
  #   b) "1 oz Foo Hops – Whirlpool at 180°F"
  #   c) "1 oz Foo Hops add with 60 minutes left in boil"
  hop_section <- str_extract(raw_text,
    "(?s)Hops Included in this Kit:[^\n]*\n(.+?)(?=\nTotal Boil Time|\\Z)")

  parse_hops <- function(keyword) {
    if (is.na(hop_section)) return(NA_character_)
    lines <- str_split(hop_section, "\n")[[1]]
    # Merge continuation lines (lines not starting with a digit)
    joined <- character()
    for (ln in lines) {
      if (str_detect(str_trim(ln), "^\\d")) {
        joined <- c(joined, str_trim(ln))
      } else if (length(joined) > 0 && str_trim(ln) != "") {
        joined[length(joined)] <- paste(joined[length(joined)], str_trim(ln))
      }
    }
    hits <- joined[str_detect(str_to_lower(joined), keyword)]
    if (length(hits) == 0) return(NA_character_)
    # Extract quantity + hop name up to the usage description
    parsed <- str_extract(hits,
      "\\d+\\.?\\d*\\s+(?:oz|LB|lb)\\s+[\\w][\\w &]+?Hops?")
    parsed <- str_trim(na.omit(parsed))
    if (length(parsed) == 0) return(NA_character_)
    paste(parsed, collapse = "\n")
  }

  bittering_hops <- parse_hops("boil")
  flavor_hops    <- parse_hops("whirlpool")
  dry_hops       <- parse_hops("dry hop")

  # 6. Original SG (take first number if range like "1.059 – 1.061")
  og_m        <- str_match(raw_text, "Estimated Original Gravity \\(OG\\):\\s*([\\d.]+)")
  original_sg <- if (!is.na(og_m[1, 2])) og_m[1, 2] else NA_character_

  # 7. Yeast
  yeast_section <- str_extract(raw_text,
    "(?s)Recommended Yeast[^:\n]*:[^\n]*\n(.+?)(?=\nNotes About|\\Z)")
  yeast_str <- if (!is.na(yeast_section)) {
    lines <- str_split(yeast_section, "\n")[[1]]
    lines <- str_trim(lines[!str_detect(lines, "Recommended Yeast|^\\s*$")])
    lines <- lines[nchar(lines) > 0]
    if (length(lines) > 0) paste(lines, collapse = "\n") else NA_character_
  } else NA_character_

  # 8. Final SG
  fg_m     <- str_match(raw_text, "Estimated Final Gravity \\(FG\\):\\s*([\\d.]+)")
  final_sg <- if (!is.na(fg_m[1, 2])) fg_m[1, 2] else NA_character_

  # 9. Flavoring ("None" or blank → NA)
  flav_section <- str_extract(raw_text,
    "(?s)Flavoring or Spices in this Kit:[^\n]*\n([^\n]+)")
  flav_line <- if (!is.na(flav_section)) {
    str_trim(str_split(flav_section, "\n")[[1]][2])
  } else NA_character_
  flavoring_str <- if (is.na(flav_line) || flav_line %in% c("", "None")) NA_character_ else flav_line

  tibble(
    recipe_name    = recipe_name,
    grains         = grains_str,
    bittering_hops = bittering_hops,
    flavor_hops    = flavor_hops,
    aroma_hops     = NA_character_,   # not differentiated in sheet format
    original_sg    = original_sg,
    yeast          = yeast_str,
    dry_hops       = dry_hops,
    final_sg       = final_sg,
    flavoring      = flavoring_str
  )
}
```

```{r}
#| label: extract-wrapper

# Auto-detect format by content (some "Instructions" files use the Sheet layout)
extract_recipe <- function(pdf_path) {
  raw_text <- paste(pdf_text(pdf_path), collapse = "\n")
  if (str_detect(raw_text, fixed("Kit Type: All Grain"))) {
    extract_recipe_instructions(pdf_path)
  } else {
    extract_recipe_sheet(pdf_path)
  }
}
```

## Test on a single recipe

```{r}
#| label: test-single

result <- extract_recipe(
  "recipe-pdfs/22nd State Amber Ale All Grain Recipe Instructions.pdf"
)

result |> t() |> as.data.frame() |> setNames("value")

result |> write_csv("data-processed/test.csv")
```

## Run on all PDFs

Once the single-recipe test looks good, run this chunk to process all PDFs and
save to CSV. Any PDFs that fail to parse will appear in the `error` column.

```{r}
#| label: run-all
#| eval: false

pdf_files <- list.files("recipe-pdfs", pattern = "\\.pdf$", full.names = TRUE)

recipes <- map(pdf_files, \(f) {
  tryCatch(
    extract_recipe(f),
    error = function(e) {
      tibble(
        recipe_name = basename(f), grains = NA, bittering_hops = NA,
        flavor_hops = NA, aroma_hops = NA, original_sg = NA,
        yeast = NA, dry_hops = NA, final_sg = NA, flavoring = NA,
        error = conditionMessage(e)
      )
    }
  )
}, .progress = TRUE) |>
  list_rbind()

write_csv(recipes, "data-processed/recipes.csv")

# Report any failures
recipes |> filter(!is.na(error))
```
