---
title: "All-grain recipe"
format: html
params:
  recipe_id: 1
---

```{r}
#| label: setup
#| warning: false
#| message: false
#| include: false

library(tidyverse)
library(janitor)
library(readxl)
library(htmltools)

recipes_raw <- read_excel("data-processed/recipes.xlsx")
pdf_lookup   <- read_csv("data-processed/recipe-pdfs.csv", show_col_types = FALSE)

recipe <- recipes_raw |> filter(id == params$recipe_id)

headline <- recipe$recipe_name

abv <- ((recipe$original_sg - recipe$final_sg) * 131.25) |> round(1)

pdf_filename <- pdf_lookup |>
  filter(id == params$recipe_id) |>
  pull(pdf_filename)
```

# `r headline`

::: callout-important

These  all-grain recipe ingredients are inserted into a standard template, so some details may be missing or included with ingredients. If you need a brewing tutorial, you might start with [Northern Brewer's guide](https://www.northernbrewer.com/blogs/new-to-brewing-start-here/how-to-brew-beer-homebrewing-101).

:::

```{r}
#| echo: false
if (!is.na(pdf_filename)) {
  pdf_path <- paste0("../recipe-pdfs/", pdf_filename)
  HTML(paste0('<p><a href="', pdf_path, '">📄 Download original recipe PDF</a></p>'))
}
```

## Measurements

| Original gravity       | Final gravity       | ABV     |
| ---------------------- | ------------------- | ------- |
| `r recipe$original_sg` | `r recipe$final_sg` | `r abv` |

## Grain bill

```{r}
#| echo: false
HTML(gsub("\r\n", "<br>", recipe$grains, fixed = TRUE))
```

### Boil additions (if any)

```{r}
#| echo: false

HTML(gsub("\r\n", "<br>", recipe$boil_additions, fixed = TRUE))
```


## Hops schedule

### Bittering

```{r}
#| echo: false

HTML(gsub("\r\n", "<br>", recipe$bittering_hops, fixed = TRUE))
```

### Flavor

```{r}
#| echo: false

HTML(gsub("\r\n", "<br>", recipe$flavor_hops, fixed = TRUE))
```

### Aroma

```{r}
#| echo: false

HTML(gsub("\r\n", "<br>", recipe$aroma_hops, fixed = TRUE))
```


## Yeast options

Possible yeasts to use with this recipe:

```{r}
#| echo: false

HTML(gsub("\r\n", "<br>", recipe$yeast, fixed = TRUE))
```


## Additions

### Dry hopping

```{r}
#| echo: false

if (is.na(recipe$dry_hops) || recipe$dry_hops == "NA") {
  HTML("No dry hop additions")
} else {
  HTML(gsub("\r\n", "<br>", recipe$dry_hops, fixed = TRUE))  
}
```

### Flavoring

```{r}
#| echo: false

if (is.na(recipe$flavoring) || recipe$flavoring == "NA") {
  HTML("No flavoring additions")
} else {
  HTML(gsub("\r\n", "<br>", recipe$flavoring, fixed = TRUE))  
}
```
